% Disciple-BD.tex
\begin{hcarentry}[section,updated]{Disciple}
\report{Ben Lippmeier}%05/16
\status{experimental, active development}
\participants{
        Ben Lippmeier, Jacob Stanley}
\makeheader

The Disciplined Disciple Compiler (DDC) is a research compiler used to investigate program transformation in the presence of computational effects. It compiles a family of strict functional core languages and supports region and effect typing. This extra information provides a handle on the operational behaviour of code that isn't available in other languages. Programs can be written in either a pure/functional or effectful/imperative style, and one of our goals is to provide both styles coherently in the same language.

\WhatsNew

DDC is in an experimental, pre-alpha state, though parts of it do work. In September this year we released DDC 0.4.3, with the following new features:

\begin{compactitem}
\item Completed desugaring of pattern alternatives.
\item Better type inference for higher ranked types, which allows explicit dictonaries
      for Functor, Applicative, Monad and friends to be written easily.
\item Automatic insertion of run and box casts is now more well baked.
\item Automatic interrogation of LLVM compiler version and generation of matching LLVM assembly syntax.
\item Added code generation for partial applications of data constructors.
\item Added support for simple type synonyms.
\end{compactitem}

\FurtherReading
  \url{http://disciple.ouroborus.net}
\end{hcarentry}
